---
title: "NBA package tutorial"
output:
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to seamlesly create plots, tables, and maps 
  for the National Biodiveristy Assessment in R with 
  the nbaR package.
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(nbaR)

tutorial_options(
  exercise.timelimit = 60,
  # A simple checker function that just returns the message in the check chunk
  exercise.checker = function(check_code, ...) {
    list(
      message = eval(parse(text = check_code)),
      correct = logical(0),
      type = "info",
      location = "append"
    )
  }
)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

In this tutorial, you will learn how to make the plots, tables, and maps used in the NBA:

-   How to make bar and donut plots of threat status, protection level, condition etc with `nba_plot()`

-   How to combine these plots into two complimentary plots of extent and number of threat status etc with `nba_plot_comb()`

-   How to plot the Red List Index using `nba_plot_RLI()`

-   How to make a bubble plot of pressures experienced by threatened species using `nba_plot_bubble()`

-   How to add a custom NBA theme to any ggplot you create using the `nba_plot_theme()` inside ggplot

-   How to make a basic table that will render to html format with `nba_tbl()`

-   How to make a basic table where the cells with categories of threat status, protection level, condition etc are coloured according to the NBA standards with `nba_tbl_colr()`

-   How to make add the a custom NBA theme to any table made with `gt()` using `nba_tbl_theme()`

-   How to make a table that cross-references the number of ecosystems or taxa by their threat status and protection level with `nba_tbl_comb()`

-   How to make a basic map of the threat status, protection level, condition etc of ecosystems using `nba_map()`

For more information on the package or any of the function please refer to it's [website](https://sanbi-nba.github.io/nbaR/)

### Setup

To practice these skills, we will use several datasets pre set in the package that you will always have access to. These datasets will allow you to test out the functions and help provide examples of how data should be formatted before reading into r. 


I've preloaded the packages for this tutorial with

```{r eval = FALSE}
library(tidyverse) 
library(nbaR)
```


## Create plots for use in the NBA

### Bar and donut plots

*`nba_plot()`*

The function creates a basic bar or donut plot with several options to customize the plot. 
You can play around with these options to test out what they do. 
The first argument is the dataframe you intend to use, so load in both examples, have a look at them and think about their structure. 


```{r nba_plot_example_data, exercise = TRUE}
NBA_example_pro_data <- NBA_example_pro_data
NBA_example_pro_data

NBA_example_thr_data <- NBA_example_thr_data
NBA_example_thr_data
```


As you can see the ecosystem function types (or taxa) are in one column, and the number of ecosystems in each threat status or protection level are in the following columns. There are two important things to note, one that the names of the columns are very important, they should follow the naming conventions of the NBA, which can be found in the list `NBA_categories`. The other is that one dataframe has a total (TOT) column and one does not. This is because you will specify which columns the function should look at, so having extra columns such as TOT does not matter, you just have to have all the threat status or protection level (or condition etc) columns. 

To make the plots you must provide:

1. **DF** The dataframe, which we now have.

2. **GROUPS** The column that contains the groups (taxa, ecosystems, ecosystem functional groups etc). This can just be the column name, and shouldn't be in inverted commas. R does not like spaces so if your column name has a space in it you can use `` these ticks to surround it to tell R that it is one name.

3. **COLs** The columns that contain the threat status (there are several ways to provide these if you are familiar with tidy select functions, but the easiest is to just provide the column positions so the number of the column that contains the first category followed by : and then the number of the column that contains the last category), for instance 2:5.

4. **CHRT** The type of plot you want, here you only have the choice between a bar or donut plot. This should be in inverted commas.

5. **NUM** Whether the values should be plotted onto the plot. There is only a TRUE or FALSE option for this. When plotting number of ecosystems or taxa this is usually preferable, but when plotting extent it is usually better to set to FALSE. The default setting is FALSE so if you don't include this argument the values won't be plotted.

6. **LAB** The label for the x axis, so either extent, number, percentage of ecosystem types for example. This should also be in inverted commas. 

7. **GRP** This is only used for the donut plots and is a TRUE or FALSE option that allows you to choose if you want to plot all the data in one donut plot or each row/group as its own donut plot.The default is FALSE so if you don't include it only one donut plot will be plotted.

6. **SAVE** If you want to automatically save a png of the plot you can add the name you would like to save it as to the save argument, otherwise you can leave it as NULL and the plot will not be saved. If you do save it you should have an outputs folder in your project to store it, if you do not you will be prompted to make one. This helps to save all plots with the same resolutions. This should also be in inverted commas.

You can test out the function in the below code chunk. 

```{r nba_plot, exercise = TRUE, exercise.eval = FALSE, warning=FALSE}
bar_plot <- nba_plot(DF = NBA_example_thr_data,
                  GROUPS = `OVERALL types`,
                  COLS = 2:5,
                 CHRT = "bar",
                 NUM = TRUE,
                 LAB = "Percentage of ecosystem types",
                SAVE = NULL)

bar_plot
```

```{r nba_plot-check}
"Good job!"
```

### Exercise 1

How could you use this same function to create a donut plot?

```{r nba_plotex4, exercise = TRUE, warning=FALSE}

```

```{r nba_plotex4-solution}
donut_plot <- nba_plot(DF = NBA_example_pro_data,
                  GROUPS = `OVERALL types`,
                  COLS = 2:5,
                 CHRT = "donut",
                 NUM = TRUE,
                 LAB = "Percentage of ecosystem types",
                 GRP = FALSE,
                SAVE = NULL)

donut_plot
```

::: {#nba_plotex4-hint}
**Hint:** Use  CHRT = "donut" and don't forget the GRP argument
:::

```{r nba_plotex4-check}
"Great work! You have mastered nba_plot!"
```


### Multi plots

*`nba_plot_comb()`*

The function creates a combined plot of two or more plots created by the `nba_plot()` function. For this function you will need an extra column that differentiates your data into the two or more plots. For example you could call the column metric, and have "extent" for the rows showing extent of ecosystems that are threatened and "number" for rows showing the number of threatened ecosystems. For the example dataset in this package I combined the NBA_example_pro_data and NBA_example_thr_data and made a column called metric that had threat_status for rows of threat status and protection_level for rows with protection level information.

You can see below how the data now looks and think about its structure. 


```{r nba_plot_comb_example_data}
NBA_example_comb_data
```


This example dataset is a bit unusual in that you will mostly be plotting two plots of threat status together, so you won't have columns of both threat status and protection level, but it will work exactly the same either way.

The nameing of the plots is automatically alphabetical and lower case.
The ordering of the plots is also automatic.
The legend is assumed to be the same and added to the bottom
of the plots.

To make the plots you must provide all the same things as for `nba_plot`:

1. **DF** The dataframe, which we now have.

2. **GROUPS** The column that contains the groups (taxa, ecosystems, ecosystem functional groups etc). This can just be the column name, and shouldn't be in inverted commas. R does not like spaces so if your column name has a space in it you can use `` these ticks to surround it to tell R that it is one name.

3. **COLs** The columns that contain the threat status (there are several ways to provide these if you are familiar with tidy select functions, but the easiest is to just provide the column positions so the number of the column that contains the first category followed by : and then the number of the column that contains the last category), for instance 2:5.

4. **CHRT** The type of plot you want, here you only have the choice between a bar or donut plot. This should be in inverted commas.

5. **NUM** Whether the values should be plotted onto the plot. There is only a TRUE or FALSE option for this. When plotting number of ecosystems or taxa this is usually preferable, but when plotting extent it is usually better to set to FALSE. The default setting is FALSE so if you don't include this argument the values won't be plotted.

6. **LAB** The label for the x axis, so either extent, number, percentage of ecosystem types for example. This should also be in inverted commas. **NB PLEASE NOTE** The label here is slightly different in that the words in the **METRICS** column will be appended onto the end of the label to differentiate the plots.

7. **GRP** This is only used for the donut plots and is a TRUE or FALSE option that allows you to choose if you want to plot all the data in one donut plot or each row/group as its own donut plot.The default is FALSE so if you don't include it only one donut plot will be plotted.

6. **SAVE** If you want to automatically save a png of the plot you can add the name you would like to save it as to the save argument, otherwise you can leave it as NULL and the plot will not be saved. If you do save it you should have an outputs folder in your project to store it, if you do not you will be prompted to make one. This helps to save all plots with the same resolutions. This should also be in inverted commas.

**Plus** there are now two extra arguments: 

7. **METRIC_COL** The column that stores the information on how to split the data.

8. **METRICS** The names you used in the metric column to split the data. These should be in inverted commas.

You can test out the function in the below code chunk. 

```{r nba_plot_comb, exercise = TRUE, exercise.eval = FALSE, warning=FALSE}
bar_plot_comb <- nba_plot_comb(
                                DF = NBA_example_comb_data,
                                GROUPS= `OVERALL types`,
                                METRIC_COL = metric,
                                METRICS = c("protection_level", "threat_status"),
                                COLS = 3:10,
                                CHRT = "bar",
                                NUM = FALSE,
                                LAB = "Percentage of ecosystem",
                                SAVE=NULL)

bar_plot_comb
```

```{r nba_plot_comb-check}
"Good job!"
```

### Exercise 2

How could you use this same function to create two donut plots?

```{r nba_plot_combex4, exercise = TRUE, warning=FALSE}

```

```{r nba_plot_combex4-solution}
bar_plot_comb <- nba_plot_comb(
                                DF = NBA_example_comb_data,
                                GROUPS= `OVERALL types`,
                                METRIC_COL = metric,
                                METRICS = c("protection_level", "threat_status"),
                                COLS = 3:10,
                                CHRT = "donut",
                                NUM = FALSE,
                                LAB = "Percentage of ecosystem",
                                GRP = FALSE,
                                SAVE=NULL)

bar_plot_comb
```

::: {#nba_plot_combex4-hint}
**Hint:** Use code from the bar plot and add CHRT = "donut" and don't forget the GRP argument
:::

```{r nba_plot_combex4-check}
"Great work! You have mastered nba_plot_comb!"
```


### RLI plots


*`nba_plot_RLI()`*

The function creates a basic Red List Index plot several options to customize the plot. 
You can play around with these options to test out what they do. 
The first argument is the dataframe you intend to use, so I load in the example data, have a look at them and think about their structure.


```{r nba_RLI_example_data}
NBA_example_RLI_data
```


As you can see the there is a year column, with the associated RLI (Red List Index) values, as well as the minimum and maximum RLI values for that year. The order of the columns does not matter as you will specify in the function which columns to use. The min and max columns are used to create creates a buffer parallel to the RLI line. You can also draw multiple lines on the same graph for multiple species if desired.

To make the plots you must provide:

1. **DF** The dataframe, which we now have

2. **YEAR** The column that contains the years

3. **RLI** The column that contains the RLI values

4. **MIN** The column that contains the minimum RLI values

5. **MAX** The column that contains the maximum RLI values

6. **GRP** The column that contains the grouping valriable. The default is NULL and if left out only a single line will be plotted 

7. **SAVE** If you want to automatically save a png of the plot you can add the name you would like to save it as to the save argument, otherwise you can leave it as NULL and the plot will not be saved. If you do save it you should have an outputs folder in your project to store it, if you do not you will be prompted to make one. This helps to save all plots with the same resolutions. This should also be in inverted commas.


### Exercise 3

How could you use this function to create an RLI plot?

```{r nba_plot_RLIex4, exercise = TRUE, warning=FALSE}

```

```{r nba_plot_RLIex4-solution}
RLI_plot <- nba_plot_RLI(DF = NBA_example_RLI_data,
                         YEAR = Years,
                         RLI = RLI,
                         MIN = min,
                         MAX = max)

RLI_plot
```

::: {#nba_plot_RLIex4-hint}
**Hint:** look at the previous example and the NBA_example_RLI_data for guidance
:::

```{r nba_plot_RLIex4-check}
"Great work! You have mastered nba_plot_RLI!"
```



### Bubble plots


*`nba_plot_bubble()`*

This function will create bubble plot intended for creation of the percentage of taxa of ecological concern impacted by various pressures. The plot created has pressures on the right and sub-pressures on the left, with taxa on the top. Bubbles are sized accourding to the perecentage of taxa of concern that are impacted by the pressure, so you must have a column with that information in your data.

The first argument is the dataframe you intend to use, so I have loaded in the example data, have a look at it and think about its structure.


```{r nba_plot_bubble_example_data}
NBA_example_bubble_data
```


Very importantly this data is all in long format. There is a column with the taxa name, and for each taxa there are several rows for each pressure and sub pressure that it experiences as well as the percentage of taxa of concern that experience that pressure. 

You should also see that some taxa have rows with pressures that they are not affected by, resulting in an NA in the last column. This is because I assume most of your data will be in wide format and when it is converted to long format these NA's appear, and they are not a problem. However, if you are creating the data from scratch in long format there is no need to include the pressures that don't affect a taxa, and you would get the following data set, which would work just as well. 

```{r nba_plot_bubble_example_dataex4, exercise = TRUE, exercise.eval = FALSE, warning=FALSE}
NBA_example_bubble_data_no_NA <- NBA_example_bubble_data %>% 
  filter(!is.na(perc_concern_under_press)) ##only keep rows where perc_concern_under_press is NOT NA

NBA_example_bubble_data_no_NA
```

To make the plots you must provide:

1. **DF** The dataframe, which we now have

2. **GROUP** The column that contains the taxa or taxa group names

3. **CAT** The column that contains overall pressure categories (named more generally CAT in case this plot is used for things other than pressures)

4. **SUB_CAT** The column that contains sub categorie/ sub pressures

5. **VALUE** The column that contains values, in this case the percentages of taxa of ecological concern that are impacted by the pressure

6. **SAVE** If you want to automatically save a png of the plot you can add the name you would like to save it as to the save argument, otherwise you can leave it as NULL and the plot will not be saved. If you do save it you should have an outputs folder in your project to store it, if you do not you will be prompted to make one. This helps to save all plots with the same resolutions. This should be in inverted commas.

**NB TO NOTE** The names of the pressures and sub-pressures are linked to their colours and must follow the standard NBA spellings and names. These can be found in `NBA_colours`

### Exercise 3

How could you use this function to create an bubble plot?

```{r nba_plot_bubbleex4, exercise = TRUE, fig.width=9, fig.height=7, warning=FALSE}

```

```{r nba_plot_bubbleex4-solution}
bubble_plot <- nba_plot_bubble(DF = NBA_example_bubble_data,
                         GROUP = taxon_group,
                         CAT = pressure,
                         SUB_CAT = sub_pressure,
                         VALUE = perc_concern_under_press,
                         SAVE = NULL)

 bubble_plot
```

::: {#nba_plot_bubbleex4-hint}
**Hint:** look at the previous example and the NBA_example_bubble_data for guidance
:::

```{r nba_plot_bubbleex4-check}
"Great work! You have mastered nba_plot_bubble!"
```



### ggplot theme


*`nba_plot_theme()`*

This function will add the default NBA aesthetics to any ggplot. Therefore if you are making a plot that is not covered by one of the previouse functions you can add the functions `nba_plot_theme()` to the end of your ggplot to format it to the correct styling

There are no arguments in this function.

### Excercise 4

Here is a point plot using ggplot

```{r nba_plot_theme_example, exercise = TRUE, warning=FALSE}

gg_plot <- NBA_example_pro_data |>
  ggplot(aes(x = `OVERALL types`, y = `Well Protected`))+
 ggplot2::geom_point()

gg_plot
```


See if you can add the NBA styling to it and see how it changes

```{r nba_plot_themeex4, exercise = TRUE, warning=FALSE}

```

```{r nba_plot_themeex4-solution}
gg_plot <- NBA_example_pro_data |>
  ggplot(aes(x = `OVERALL types`, y = `Well Protected`))+
 ggplot2::geom_point()+
  nba_plot_theme()

gg_plot
```

::: {#nba_plot_themeex4-hint}
**Hint:** look at the nba_plot_theme function
:::

```{r nba_plot_themeex4-check}
"Great work! You have mastered nba_plot_theme!"
```


##  Create tables for use in the NBA

The NBA makes use of several tables. So far we only have functions to make a basci table with the basic NBA styling. This table is rendered through the `nba_tbl()` function and was made to be rendered in quarto to an html format. Therefore when you render the basic table here it does not look right, when rendered in html the sizing becomes correct. There is also an option to create a table where you can colour a column by its values `nba_tbl_col()`, as long as the values are one of the categories in `NBA_colours`. 

There is then a function to create the table of protection level of threatened ecosystems `nba_tbl_comb()` and an NBA theme function `nba_tbl_theme()` that can be added onto any gt() table to give it the offcial NBA style.

For the tables we can use the same example datasets as we did for the plots. 


### Basic tables

*`nba_tbl()`*

The function creates a basic table with a purple heading bar. We will be using these two examples again:


```{r nba_tbl_example_data, exercise = TRUE}

NBA_example_pro_data

NBA_example_thr_data
```

The table function only takes the dataframe and styles it like a table. Therefore the table will have the exact same structure as the dataframe, and any structure you want the table to have, you need to do to the dataframe before using the `nba_tbl()` function. 

To make the table you must provide:

1. **DF** The dataframe, which we now have.


You can test out the function in the below code chunk. 

```{r nba_tbl, exercise = TRUE, exercise.eval = FALSE, warning=FALSE}
tbl <- nba_tbl(NBA_example_pro_data)

tbl
```

```{r nba_tbl-check}
"Good job!"
```

### Exercise 5

Can you create a quarto document, load nbaR, run the table function on some data in a code chunk and knit it to an html format to see what the table will look like on a website?




### Tables with cells coloured according to their values

*`nba_tbl_colr()`*

A basic table with a purple heading bar and coloured blocks around the categories. Therefore for this table the categories are not the column headings, but now they are values in a column of protection level or threat status etc categories.

You can see below how the data now looks and think about its structure. 


```{r nba_plot_col_example_data}
tbl <- NBA_example_pro_data %>%
  pivot_longer(2:5, names_to = "protection_level")

tbl
```

To then colour the protection level values you can use the `nba_tbl_colr()` function. 

For the function you just need:


1. **DF** The dataframe, which we now have.

2. **COL** The name of the column where the categories are



### Excercise 6


See if you can change the NBA_example_pro_data to the correct format and make it into a table where the protection levels are coloured according to their values

```{r nba_tbl_colrex4, exercise = TRUE, warning=FALSE}

```

```{r nba_tbl_colrex4-solution}
tbl <- NBA_example_pro_data %>%
  pivot_longer(2:5, names_to = "protection_level") %>% 
  nba_tbl_colr(protection_level)

tbl
```

::: {#nba_tbl_colrex4-hint}
**Hint:** Use the code from the example data chunk above.
:::

```{r nba_tbl_colrex4-check}
"Great work! You have mastered nba_tbl_colr!"
```



### Table of protection level of threatened ecosystems

*`nba_tbl_comb()`*

A function to style a table with threat status on the horizontal rows and protection level on the vertical columns, with number of ecosystems or taxa that share those categories in the cells, with total çolumns added.

For this data I have taken the marine pelagic ecosystem map (which is currently a GeoPackage, which we now use instead of shapefiles). To create this table from a map of ecosystems, you just need to have the threat status and protection level of the ecosystems as fields/attributes in your map shp/gpkg file and to remove the polygons from the dataset. ALternatively you can make a csv/ excel file of your protection level of threatened ecosystems and just read that into r. 

We use the **sf** package for all our spatial analysis with vectors and I highly recommend you do to, it allows you to treat a spatial object like a normal dataframe which makes your life incredibly easy. 

In the below code you can see how an sf map object (that you would have read into R) would look like and think about its structure. 


```{r nba_tbl_comb_example_data}
NBA_example_map_data


```



For the function you just need:


1. **DF** The dataframe, which we now have.

2. **GROUP** The name of the column where the ecosystem or taxa names are stored

3. **THR** The name of the column where the threat status of the ecosystems/ taxa are stored

4. **PRO** The name of the column where the protection level of the ecosystems/ taxa are stored

5. **FILE** States if you inital DF file is a spatial file, or a csv by stating FILE = "spatial" or file = "csv"

**NB TO NOTE** The threat status and protection level categories must again match the spelling and conventions in the `NBA_colours` and `NBA_categories` lists for the colours to map correctly.

Also if you start from a spatial file you will need to add the DF, GROUP, THR, and PRO, and FILE arguments, but if you have already fromatted the data and read it in from a csv then you just need to provide the DF, THR and FILE arguments.



### Excercise 7


See if you can change the NBA_example_map_data to the correct format and make it into a table of protection level of threatened ecosystems

```{r nba_tbl_combex4, exercise = TRUE, warning=FALSE}

```

```{r nba_tbl_combex4-solution}
thr_pro_tbl <- NBA_example_map_data 
nba_tbl_comb(GROUP = P_EcosysType,
THR = threat_status,
PRO = protection_level, 
FILE = "spatial")

thr_pro_tbl
```

::: {#nba_tbl_combex4-hint}
**Hint:** Use the code from the example data chunk above.
:::

```{r nba_tbl_combex4-check}
"Great work! You have mastered nba_tbl_comb!"
```


### Table theme to be added to any gt table object

*`nba_tbl_theme()`*

Function to apply gt styling to a table.

To make tables in R there are two main packages we use, Kable (which is actually part of a family of functions kable, knitr, and kableExtra that all work together) and gt. They are both powerful packages that can create complex tables. Kable has been around longer and so has slightly more help and functions, but both are capable of creating great tables. Which you use is completely up to your preference, and we use both in this package. 

If you do use gt to create a gt object as your table, you can use the `nba_tbl_theme()` function at the end to style your table to match the official NBA style.

For this function you can use any of the example datastes, except for the map which you would have to first remove the polygon 'geom' column as in the previous example. I show an example below of how to create a basic gt table from one of the example datasets.


```{r nba_tbl_theme_example_data}
gt_tbl <- NBA_example_thr_data %>%
  gt::gt() 

gt_tbl
```



For the function you just need:


1. **DF** The dataframe, which we now have



### Excercise 8


See if you can take make the gt table object as in the previous chunk and add the NBA styling to it

```{r nba_tbl_themeex4, exercise = TRUE, warning=FALSE}

```

```{r nba_tbl_themeex4-solution}
gt_tbl <- NBA_example_thr_data %>%
  gt::gt() %>%
 nba_tbl_theme()

gt_tbl
```

::: {#nba_tbl_themeex4-hint}
**Hint:** Use the code from the example data chunk above and pipe in the function or add the object inside the function
:::

```{r nba_tbl_themeex4-check}
"Great work! You have mastered nba_tbl_theme!"
```



##  Create maps  for use in the NBA

The NBA makes use of several maps, namely their ecosystem maps, maps of protection level, threat status, condition, pressures etc. So far in the package we have made a function to create the protection level, threat status and condition maps, and hopefully will be able to add the others in future. This function is called `nba_map()`.

For the map we can use the example dataset `NBA_example_map_data`.


### Map function

*`nba_map()`*

The function creates a basic map of ecosystems coloured by categories such as protection level etc.

Below is an example dataset. To get a map dataset in R you can get by reading in a GeoPackage or shapefile using the sf package. **NB TO NOTE** These protection level and threat status categories are completely arbitrary and not real values for these ecosystems. 

The function can take a dataset with either one polygon or multipolygon per ecosystem (each row has a unique ecosystem type) or multiple rows per ecosystem type. This example dataset has already been concatenated to have only one row per ecosystem type, but most maps especially the ecosystem maps will not come like this, just know it's not a problem and the function can deal with these datasets.

```{r nba_map_example_data, exercise = TRUE}

NBA_example_map_data
```


To make the map you must provide:

1. **DF** The dataframe, which we now have

2. **COLS** The columns to use to group the ecosystems by when combining all the individual polygons of an ecosystem type into one multipolygon (concatenating the dataset to one row per ecosystem type). This is set to NULL by default so if you already have one ecosystem type you can just leave it out. Otherwise it accepts names using c(ecosystem_name, threat_status) to select the columns you want to group by (and by default keep).

2. **GEOM** The name of the column with the geometries or polygons

3. **CAP** The figure capetion to go at the bottom of the plot

4. **FILL** The column with the values/categories you would like to colour the plot by. These categories must follow the naming and spelling conventions of the NBA, which can be seen in the `NBA_colours` and `NBA_categories` lists.


You can test out the function in the below code chunk. 

```{r nba_map, exercise = TRUE, exercise.eval = FALSE, warning=FALSE}
map <- nba_map(DF = NBA_example_map_data,
              GEOM = geometry,
              CAP = "Figure 1. A map of the protection level of marine pelagic ecosystem types in South Africa",
            FILL = protection_level)

map
```

```{r nba_map-check}
"Good job!"
```


### Exercise 9

See if you can create a map of threat status of ecosystems

```{r nba_mapex4, exercise = TRUE, warning=FALSE}

```

```{r nba_mapex4-solution}
map <- nba_map(DF = NBA_example_map_data,
              GEOM = geometry,
              CAP = "Figure 1. A map of the threat status of marine pelagic ecosystem types in South Africa",
            FILL = threat_status)

map
```

::: {#nba_mapex4-hint}
**Hint:** Use the code from the example data chunk above, change the FILL argument, and remember to change the caption in the CAP argument.
:::

```{r nba_mapex4-check}
"Great work! You have mastered nba_map!"
```



